\documentclass[subfile]{main.tex}


\tableofcontents

\section{Introduction}
In this paper, we intend to discuss about certain properties of the ellipses outside of the cartesian plane. We use synthetic properties to arrive at theorems regarding ellipses, some of which would be just special cases of properties of conics in general, which are just being applied to ellipses to satisfy the purpose of this paper. It is highly recommended to also read about their extensions in other conics as well. The sources can be found in the references section.

\subsection{Buffer Problems}
In this subsect, I dedicate to trash all sorts of problems I have ever encountered from known sources, which I would be taking out later on from the publish but it acts as a buffer to aid in explanation of the other harder topics coming fourth

\begin{problem}[AIME 1985]
An ellipse has foci at $(9,20)$ and $(49,55)$ in the $xy$-plane and is tangent to the $x$-axis. What is the length of its major axis?
\end{problem}

\begin{problem}[abacadaea's mock AIME]
The point $(10,26)$ is a focus of a non-degenerate ellipse tangent to the positive $x$ and $y$ axes. The locus of the center of the ellipse lies along graph of $ax-by+c=0,$ where $a,b,c$ are positive integers with no common factor other than $1$. Find $a+b+c.$
\end{problem}

\begin{problem}[PUMaC 2012 A4/B6]
A square is inscribed in an ellipse such that two sides of the square respectively pass through the two foci of the ellipse. The square has a side length of $4$. The square of the length of the minor axis of the ellipse can be written in the form $a + b\sqrt{c}$ where $a, b$, and $c$ are integers, and $c$ is not divisible by the square of any prime. Find the sum $a + b + c$.
\end{problem}

\begin{problem}[PUMaC 2020 A6/B8]
Triangle $ABC$ has side lengths $13$, $14$, and $15$. Let $E$ be the ellipse that encloses the smallest area which passes through $A, B$, and $C$. The area of $E$ is of the form $\frac{a \sqrt{b}\pi}{c}$ , where $a$ and $c$ are coprime and $b$ has no square factors. Find $a + b + c$.
\end{problem}
